\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{pgfplots}
\usepackage[left=1.5cm,right=1.5cm,top=2cm,bottom=2cm]{geometry}
\begin{document}
\section*{Poisson equation in 2D}
We want to solve the two-dimensional Poisson equation numerically:
\begin{equation}
(\partial_{xx} + \partial_{yy}) u = f,
\end{equation}
where $u$ and $f$ are spectral coefficients. 

\subsection*{Implementation}
Multivariate problems can be described well with Kronecker notation, i.e.
\begin{equation}
(D_2 \otimes I + I \otimes D_2) u = f,
\end{equation}
where $D_2$ is the discrete differentiation operator, $I$ the identity matrix and $u$ and $f$ are the flattened one-dimensional vectors.
We impose boundary conditions by using basis recombinations $u=(S\otimes S)v$, where $v$ are spectral coefficients in a function base that already satisfies the boundary conditions. Here we will impose Dirichlet BCs on all sides. In this case $u\in\mathbb{R}^{N\times N}$, $v\in\mathbb{R}^{N-2\times N-2}$ and $S\in\mathbb{R}^{N\times N-2}$. We get
\begin{equation}
(D_2 S \otimes S + S \otimes D_2 S) v = f.
\label{eq:3}
\end{equation}
Our intent is to decouble the $x$ and $y$ dimension, which allows the use of efficient solvers. First, we multiply \eqref{eq:3} with $S^{-1}$ from the left
\begin{equation}
(S^{-1} D_2 S \otimes S + I \otimes D_2 S) v = (S^{-1}  \otimes I)f.
\label{eq:4}
\end{equation}
Applying the eigendecomposition $S^{-1} D_2 S = Q \Lambda Q^{-1}$
\begin{equation}
(Q \Lambda Q^{-1} \otimes S +  I \otimes D_2 S) v = (S^{-1}  \otimes I)f.
\label{eq:5}
\end{equation}
and multiplying with $Q^{-1}$ from the left
\begin{equation}
(\Lambda Q^{-1} \otimes S +  Q^{-1} \otimes D_2 S) v = (Q^{-1} S^{-1}´  \otimes I)f
\label{eq:6}
\end{equation}
or simplified
\begin{equation}
(\Lambda \otimes S +   I \otimes D_2 S) \widehat{v} = \widehat{f}
\label{eq:7}
\end{equation}
with $(Q^{-1} \otimes I) v = \widehat{v}$ and $(Q^{-1} S^{-1}  \otimes I)f = \widehat{f}$. Since $I$ and $\Lambda$ are both diagonal matrices, eq. \eqref{eq:7} decoubles in $x$ and $y$, such that we get a couple of one-dimensional equations
\begin{equation}
(\lambda_i S + D_2 S) \widehat{v}_i = \widehat{f}_i,
\label{eq:8}
\end{equation}
where $i$ runs from $0$ to $N-3$.\\

\textbf{In short:}\\

\textit{Preprocessing}
\begin{itemize}
\item Eigendecomposition $S^{-1} D_2 S = Q \Lambda Q^{-1}$
\end{itemize}

\textit{Main}
\begin{itemize}
\item (i) Transform $\widehat{f} = (Q^{-1} S^{-1} \otimes I)f$
\item (ii) Solve $(\lambda_i  S +  D_2 S) \widehat{v}_i = \widehat{f}_i$, where $i=0,..,N-3$.
\item (iii) Transform $v = (Q \otimes I) \widehat{v}$.
\end{itemize}

\textbf{Notes:}\\

Step (ii) can be preconditioned with $D_2^{-1}$, i.e.
\begin{itemize}
\item (ii) Solve $(\lambda_i  D_2^{-1}S +  S) \widehat{v}_i = D_2^{-1}\widehat{f}_i$
\end{itemize}

which is banded, instead of upper triangular.

\subsection*{Example}
See \textit{examples/poisson2d\_cheb\_dirichlet.rs}.

%\subsection*{Convergence}
%\begin{tikzpicture}
%    \begin{semilogyaxis}[
%    xlabel=$N$,
%    ylabel=$\left\lVert v-v_t\right\rVert/\left\lVert v_t\right\rVert$,
%    ]
%        %\addplot[mark=o, mark size=2pt, black, thick] coordinates {(8,2.07199e-7) (16,2.03066e-17) (32,2.71877e-13) (64,4.34132e-7) (128,1.23877e-1)};
%
%    \end{semilogyaxis}
%\end{tikzpicture}
\end{document}